A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2018-01-30 paper.tex
saving

%%% General usepackages, definitions, etc.
\usepackage[acronym,toc]{glossaries}
\usepackage{graphicx}
\usepackage{booktabs} % nice rules for tables
\usepackage{microtype} % if using PDF
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{moreverb} % for verbatim snippets of code
\usepackage{fancyvrb}
\usepackage{tabularx} % for tables with line breaks
\usepackage{threeparttable} % for tables with notes
\usepackage[capitalize, noabbrev]{cleveref} % for reference multiple figures
\usepackage{calc} % allows for arithmetic on latex variables
\usepackage{float} % allows for figures to be placed explicitly
\usepackage{algorithm2e} % for algorithms
\usepackage{subfig}
\usepackage{multirow} % combining rows in tables
\usepackage{footnote}
\usepackage{titlesec} % for using \titleformat
\usepackage{fixltx2e} % for using \textsubscript
\usepackage{bashful}
\usepackage{xspace}
\usepackage{color}
%%%%%

@@

@@

%%% Acronyms
\newacronym{ssps}{SSPs}{Shared Socioeconomic Pathways}
\newacronym{rcps}{RCPs}{Representative Concentration Pathways}
\newacronym{cmip6}{CMIP6}{Coupled Model Intercomparison Project (Phase 6)}
\newacronym{smip}{ScenarioMIP}{Scenario Model Intercomparison Project}
\newacronym{ceds}{CEDS}{Community Emissions Data System}
\newacronym{iam}{IAM}{Integrated Assessment Model}
\newacronym{esm}{ESM}{Earth System Model}
\newacronym{ghg}{GHG}{Greenhouse Gas}
\newacronym{luc}{LUC}{Land-use Change}
\newacronym{luluc}{LULUC}{Land-use and Land-use Change}
\newacronym{weu}{WEU}{Western Europe}
\newacronym{nam}{NAM}{North America}
\newacronym{pao}{PAO}{Pacific OECD}
\newacronym{eeu}{EEU}{Central and Eastern Europe}
\newacronym{fsu}{FSU}{Former Soviet Union}
\newacronym{pas}{PAS}{Other Pacific Asia}
\newacronym{cpa}{CPA}{Centrally-planned Asia and China}
\newacronym{sas}{SAS}{South Asia}
\newacronym{mea}{MEA}{Middle East and North Africa}
\newacronym{afr}{AFR}{Sub-Saharan Africa}
\newacronym{lam}{LAM}{Latin America and the Caribbean}

@@

\subsection{\code{aneris} Workflow}\label{sec:workflow}

The full \code{aneris} workflow is comprised of a number of components shown
graphically in Figure \ref{fig:workflow}. Unharmonized model data and a
run-control configuration are read in via an Excel spreadsheet. Data is assumed
to be in the IAMC format, i.e., using \code{Model}, \code{Scenario},
\code{Region}, \code{Variable}, and \code{Unit} columns in addition to columns
representing each modeled time period.

Users are able to control the harmonization process via a number of options. The
primary mechanism by which users control the process is by providing
\textit{override} methods for any combination of region and variable (i.e.,
sector and gas species).  In practice, it may be possible that not all default
methods chosen will provide robust harmonized trajectories, especially if there
is a significant difference between historical and model values in the
harmonization year, if there is significant upward or downward movement in the
model trajectory, or if there are known discrepancies in sectoral definition
between the IAM and historical data source. In such cases, override methods can
ameliorate the issues associated with the default method choice. In order to
help identify these cases, harmonization \textit{diagnostics} are provided which
analyze the relative difference between harmonized and unharmonized trajectories
at their mid (if $> 400$\%) and end-points (if $> 200$\%). If override methods
are provided, they are used instead of the default methods as determined by the
default method decision tree. Furthermore, users can set the above-mentioned
thresholds as well as the LUC method used in the decision tree (see Figure
\ref{fig:decision_tree}). Further detail of input parameters can be found
\href{http://mattgidden.com/aneris/config.html}{online}.

\begin{figure}
  \begin{center}
    \includegraphics[height=0.9\textheight]{aneris-workflow.pdf}
    \caption[]{
      \label{fig:workflow}
      The full harmonization process as executed by \code{aneris}. Operations
      that can be configured with user-based input configurations are shown in
      purple. The core harmonization process is shown in yellow.  }
  \end{center}
\end{figure}

>>>>

%%% General usepackages, definitions, etc.
\usepackage[acronym,toc]{glossaries}
\usepackage{graphicx}
\usepackage{booktabs} % nice rules for tables
\usepackage{microtype} % if using PDF
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{moreverb} % for verbatim snippets of code
\usepackage{fancyvrb}
\usepackage{tabularx} % for tables with line breaks
\usepackage{threeparttable} % for tables with notes
\usepackage[capitalize, noabbrev]{cleveref} % for reference multiple figures
\usepackage{calc} % allows for arithmetic on latex variables
\usepackage{float} % allows for figures to be placed explicitly
\usepackage{algorithm2e} % for algorithms
\usepackage{multirow} % combining rows in tables
\usepackage{footnote}
\usepackage{titlesec} % for using \titleformat
\usepackage{fixltx2e} % for using \textsubscript
\usepackage{bashful}
\usepackage{xspace}
\usepackage{color}
\usepackage{subfig}
\usepackage{minted}
%%%%%

@@

%%%python syntax highlighting
% Default fixed font does not support bold face
\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{10} % for bold
\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{10}  % for normal

% Custom colors
\usepackage{color}
\definecolor{deepblue}{rgb}{0,0,0.5}
\definecolor{deepred}{rgb}{0.6,0,0}
\definecolor{deepgreen}{rgb}{0,0.5,0}

\usepackage{listings}

% Python style for highlighting
\newcommand\pythonstyle{\lstset{
language=Python,
basicstyle=\ttm,
otherkeywords={self},             % Add keywords here
keywordstyle=\ttb\color{deepblue},
emph={MyClass,__init__},          % Custom highlighting
emphstyle=\ttb\color{deepred},    % Custom highlighting style
stringstyle=\color{deepgreen},
frame=tb,                         % Any extra options here
showstringspaces=false            % 
}}

% Python environment
\lstnewenvironment{python}[1][]
{
\pythonstyle
\lstset{#1}
}
{}

@@

%%% Acronyms
\newacronym{ssps}{SSPs}{Shared Socioeconomic Pathways}
\newacronym{rcps}{RCPs}{Representative Concentration Pathways}
\newacronym{cmip6}{CMIP6}{Coupled Model Intercomparison Project (Phase 6)}
\newacronym{smip}{ScenarioMIP}{Scenario Model Intercomparison Project}
\newacronym{ceds}{CEDS}{Community Emissions Data System}
\newacronym{iam}{IAM}{Integrated Assessment Model}
\newacronym{esm}{ESM}{Earth System Model}
\newacronym{ghg}{GHG}{Greenhouse Gas}
\newacronym{luc}{LUC}{Land-use Change}
\newacronym{luluc}{LULUC}{Land-use and Land-use Change}
\newacronym{weu}{WEU}{Western Europe}
\newacronym{nam}{NAM}{North America}
\newacronym{pao}{PAO}{Pacific OECD}
\newacronym{eeu}{EEU}{Central and Eastern Europe}
\newacronym{fsu}{FSU}{Former Soviet Union}
\newacronym{pas}{PAS}{Other Pacific Asia}
\newacronym{cpa}{CPA}{Centrally-planned Asia and China}
\newacronym{sas}{SAS}{South Asia}
\newacronym{mea}{MEA}{Middle East and North Africa}
\newacronym{afr}{AFR}{Sub-Saharan Africa}
\newacronym{lam}{LAM}{Latin America and the Caribbean}
\newacronym{cli}{CLI}{Command Line Interface}

@@

\subsection{\code{aneris} Python Implementation and Workflow}\label{sec:workflow}

We herein present \code{aneris}' Python implementation and conceptual
workflow. The implementation is depicted in Figure \ref{fig:software}, and the
conceptual workflow is depicted in Figure \ref{fig:workflow}. The software is
composed of three primary components: the \texttt{HarmonizationDriver},
\texttt{Harmonizer}, and data processing routines (shown in red, green, and
blue, respectively in Figure \ref{fig:software-workflow}).

The \texttt{HarmonizationDriver} is an object designed to interface with
user-provided data and configuration files. Input data is assumed to be an Excel
file the in the in the IAMC format, i.e., using \code{Model}, \code{Scenario},
\code{Region}, \code{Variable}, and \code{Unit} columns in addition to columns
representing each modeled time period, which is the standard format within the
IAM community. It is responsible for downselecting data into unique combinations
for each model and scenario and then invoking the harmonization
process. Finally, it adds any exogenous emissions trajectories as required by
the user and recompiles the results and writes them to disc. The
\texttt{HarmonizationDriver} acts the primary interface for high-level users as
shown in Listing \ref{lst:tutorial}. Furthermore, a \gls{cli} is provided

\begin{python}[caption=High-level user interaction with the \texttt{HarmonizationDriver} taken from the online \href{http://mattgidden.com/aneris/config.html}{tutorial}, label=lst:tutorial]
from aneris.tutorial import load_data
model, hist, driver = load_data()
for scenario in driver.scenarios():
    driver.harmonize(scenario)
harmonized, metadata = driver.harmonized_results()
\end{python}

\begin{lstlisting}[language=bash, caption=bash, label=lst:cli]
$ aneris -h
usage: aneris [-h] [--history HISTORY] [--regions REGIONS] [--rc RC]
              [--output_path OUTPUT_PATH] [--output_prefix OUTPUT_PREFIX]
              input_file
    Harmonize historical trajectories to data in the IAMC template format.
    Example usage:
    aneris input.xlsx --history history.csv --regions regions.csv
positional arguments:
  input_file            Input data file.
optional arguments:
  -h, --help            show this help message and exit
  --history HISTORY     Historical emissions in the base year.
  --regions REGIONS     Mapping of country iso-codes to native regions.
  --rc RC               Runcontrol YAML file (see
                        http://mattgidden.com/aneris/config.html for
                        examples).
  --output_path OUTPUT_PATH
                        Path to use for output file names.
  --output_prefix OUTPUT_PREFIX
                        Prefix to use for output file names.
\end{lstlisting}

The \texttt{Harmonizer} is a Python class whose concern is to harmonize various
model value trajectories given historical data and possible user method
\textit{overrides}, i.e., non-default methods. It is used by the
\texttt{HarmonizationDriver}; however it is also available to the user as a
first-class object. The \texttt{Harmonizer} requires that input data conform to
the \code{aneris} calculation data format, which expliciltly sepearates the
emissions species from the sector contributing the emissions (these are combined
in the single \code{Variable} column in the IAMC format). Because the
\texttt{Harmonizer} is configure for a single instance of a model and scenario,
the canonical data format includes columns named \texttt{region},
\texttt{sector}, \texttt{gas}, and \texttt{units}. Once configured with
appropriate input data (model and history) as well as potential method
overrides, the \texttt{Harmonizer}'s \texttt{harmonize()} method can be invoked
which returns a \texttt{pandas.DataFrame} of harmonized data. The object can
additionally be queried directly as to its \texttt{default\_methods()},
\texttt{methods()} (i.e., methods used with overrides), and \texttt{metadata()}
(i.e., methods used with all branching information along each path in the
decision tree).

There are also a variety of tools and utilities provided to users and also used
by the \texttt{HarmonizationDriver} in order to process both input and output
data. These include an \texttt{EmissionsAggregator} class and related routines
used to combine historical emissions to native model regions (where historical
data is defined at a higher spatial resolution than a model; see, e.g., Figure
\ref{fig:regions}) and a \texttt{FormatTranslator} class which defines an
interface for translating \texttt{pandas.DataFrame}s between the IAMC format
expected for input and output data and the calculation format used by
\code{aneris}' \texttt{Harmonizer}.

\begin{figure}
  \subfloat[Python Implementation\label{fig:software}]{
    \includegraphics[width=0.775\textwidth]{implementation_schematic.pdf}
    }
  \subfloat[Workflow\label{fig:workflow}]{
    \includegraphics[width=0.225\textwidth]{aneris-workflow.pdf}
    }
    \caption{
      The full harmonization process as executed by \code{aneris}. Operations
      that can be configured with user-based input configurations are shown in
      purple. The core harmonization process is shown in yellow.  }
    \label{fig:software-workflow}
\end{figure}

Users are able to control the harmonization process via a number of options
(with examples provided
\href{http://mattgidden.com/aneris/config.html}{online}). The primary mechanism
by which users control the process is by providing \textit{override} methods for
any combination of region and variable (i.e., sector and gas species).  In
practice, it may be possible that not all default methods chosen will provide
robust harmonized trajectories, especially if there is a significant difference
between historical and model values in the harmonization year, if there is
significant upward or downward movement in the model trajectory, or if there are
known discrepancies in sectoral definition between the IAM and historical data
source. In such cases, override methods can ameliorate the issues associated
with the default method choice. In order to help identify these cases,
harmonization \textit{diagnostics} are provided which analyze the relative
difference between harmonized and unharmonized trajectories at their mid (if $>
400$\%) and end-points (if $> 200$\%). If override methods are provided, they
are used instead of the default methods as determined by the default method
decision tree. Furthermore, users can set the above-mentioned thresholds as well
as the LUC method used in the decision tree (see Figure
\ref{fig:decision_tree}).
